% Do not chage this part
\begin{center}
  \textbf{{\Large ACKNOWLEDGEMENT}}\\[60pt]
\end{center}
\addcontentsline{toc}{chapter}{\textbf{\normalsize{\emph{ACKNOWLEDGEMENT}}}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% You need to change the text for acknowledgement
First and foremost, I express my deepest gratitude to \textbf{Almighty Allah} for bestowing His blessings on me and enabled me to successfully accomplish this work. 

I would like to express my sincere appreciation and gratitude to my supervisor \textbf{Dr.\ Hossen Asiful Mustafa}, who is a great mentor leading me to the research field of computer vision and deep learning. His cooperation and guidance has been a constant source of encouragement in shaping this work throughout my master thesis period. Without his great help and invaluable suggestions I would hardly finish this thesis. 

I will always appreciate my classmates and brothers \textbf{Md Abdul Aowal, Abu Noman}, and \textbf{Imran Khan} for their firm-backing and support that has greatly helped me in accomplishing the research work. I am also highly grateful for the support of my roommate cum brother \textbf{Abdullah Al Mahmud}, who always tolerated my frustration. We studied together and shared a lot of discussion which were very helpful for this research.

Finally, I want to dedicate the essence of my purest respect to my parents and to my colleagues for providing me with  support and continuous inspiration throughout my years of study and through the process of writing this thesis. This accomplishment would not have been possible without them. Thank you.

\vspace*{20.0mm}

\begin{minipage}[t]{0.2\textwidth}
  Dhaka\par
  \thesisdate
\end{minipage}%
\hfill
\begin{minipage}[t]{0.45\textwidth}
  \begin{enumerate}
    \vspace{-0.75\baselineskip}
    \DTLforeach{ThesisStudents}{\StudentName=Column1}{\item[]\StudentName}
  \end{enumerate}
\end{minipage}

\endinput
