% Do not chage this part
\begin{center}
  \textbf{{\Large Acknowledgement}}\\[60pt]
\end{center}
\addcontentsline{toc}{chapter}{\textbf{Acknowledgement}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% You need to change the text for acknowledgement
First and foremost, I express my deepest gratitude to \textbf{Almighty Allah} for bestowing His blessings on me and giving me the ability to accomplish this work successfully. 

I would like to express my deepest sense of thankfulness and gratitude to my thesis supervisor \textbf{Dr.\ Hossen Asiful Mustafa}, Assistant Professor, IICT, BUET for leading me into the research field of computer vision and deep learning. His scholarly guidance, constant and energetic supervision and valuable advice made this work a successful one. He has been a continuous source of inspiration and a real motivating force throughout my research work. Also, I am also extremely grateful to him for providing me a high-end GPU instance to accomplish this work. 

I would like to thank my classmates \textbf{Md Abdul Aowal, Abu Noman}, and \textbf{Imran Khan} for their firm-backing and co-operation. I am truly grateful to my roommate cum brother \textbf{Abdullah Al Mahmud} for his endless support, and encouragement with my studies and works. He always tolerates my frustration. We studied together and shared a lot of discussions which were very helpful for this research.

Finally, I want to dedicate the essence of my purest respect to my parents and to my colleagues for providing me with support throughout my years of study and through the process of writing this thesis. This accomplishment would not have been possible without them. Thank you.

\vspace*{20.0mm}

\begin{minipage}[t]{0.2\textwidth}
  Dhaka\par
  \thesisdate
\end{minipage}%
\hfill
\begin{minipage}[t]{0.45\textwidth}
    \vspace{-0.75\baselineskip}
    \author
\end{minipage}

\endinput
